---
title: "Model Building and Selection Case (Using NMB Pretended Dataset)"
author: "Laban Masunga"
date: '`r Sys.Date()`'
output: html_document
---

##	Dataset Description and the intended outcome  
A sports science experiment was conducted at NMB Bank where 6 randomly choose members of staff each supervised 5 different randomly selected participants to see if they could complete an exercise programme or not. On different occasions the participant undertook an exercise programme with different intensities and of different lengths. Half the participants had undertaken exercise training in the previous weeks and the others had not.
The nmb_miniproject.txt (text file) (available via  provides the data, which includes the following variables:
	
	Staff -  i is the unique ID for the i^th staff member
	
	Participant  - the j^th participant for the i^th member staff
	
	complete - whether the participant completed the programme (1) or did not complete the program (0)
	
	intensity - measure of the intensity of the programme (higher values are more intense)
	
	length - length in minutes of the exercise programme
	
	training - whether the participant took part in training (1) or did not take part in training (0)

This mini-project will look at fitting appropriate models, what can be concluded about the effect of the staff, participant, intensity, length and training on whether the participant completed the programme. 

The primary areas to be covered will be:

	(i) Fitting appropriate models to the data; 
	
	(ii) Using appropriate model selection techniques;
	
	(iii) Drawing appropriate conclusions, aimed at a data scientist manager. 



### Setting your directory and importing dataset
setwd("C:/OFFICE_/NMB_quick_miniproject")

```{r setup, include=FALSE}
nmbdata<-read.delim("nmb_miniproject.txt", head=T)
#knitr::opts_chunk$set(echo = TRUE)
```

### R Data Exploration

This section will give simple data exploration

```{r explore}
summary(nmbdata)
```

### Doing simple checks, eg. dimension of dataset and any suspected missing values 
```{r data check}
dim(nmbdata)
sum(is.na(nmbdata))

```

### Including Simple Plots

You can also embed plots, for example:

```{r plot, echo=T}
library(ggplot2)
ggplot(data = nmbdata, mapping = aes(x = complete, y = length))+geom_point()
ggplot(data = nmbdata, mapping = aes(x = complete, y = intensity))+geom_point()
```

### Doing some statistical analysis


```{r data, echo=T}
library(lme4)
modelilt <- glm(formula=complete~ intensity+length+factor(training),data = nmbdata,family = binomial())
summary(modelilt)
```

Note that if the `echo` parameter will be changed to 'FALSE' or 'F', the code chunk will prevent printing of the R code that generated the plot.


# Main Analysis

Using the t-test both Intensity and training have p-values < 0.05, so have covariate values significantly different zero. Suggesting these covariates should be included in the model. However, as the p-value for length is 0.74, this not significantly different from zero, so should not be included in the model.

### R code for fitting a selection of other models:
```{r fitting a selection of other models, echo=T}
 modelil <- glm(formula=complete~ intensity+length,data = nmbdata,family = binomial())
 modelit <- glm(formula=complete~ intensity+factor(training),data = nmbdata,family = binomial())
 modellt <- glm(formula=complete~ length+factor(training),data = nmbdata,family = binomial())
modeli <- glm(formula=complete~ intensity,data = nmbdata,family = binomial())
 modell <- glm(formula=complete~ length,data = nmbdata,family = binomial())
 modelt <- glm(formula=complete~ factor(training),data = nmbdata,family = binomial())
 modelconst <- glm(formula=complete~ 1,data = nmbdata,family = binomial())
```


### AIC model selection:
```{r, AIC Model selection}
library(lmerTest)
 AICvalues <- c(AIC(modelilt),AIC(modelil),AIC(modelit),AIC(modellt),AIC(modeli),AIC(modell),AIC(modelt), AIC(modelconst))
 modelname <- c("intensity+length+training", "intensity+length", "intensity+training","length+training","intensity","length","training","const")
 AICvaluesr <- round(AICvalues,digits=2)
 deltaAIC <-AICvalues-min(AICvalues)
 deltaAICr <-round(deltaAIC,digits=2)
 AICweight <- round(exp(-deltaAIC/2)/sum(exp(-deltaAIC/2)),digits=2)
 AICtable <- data.frame(modelname,AICvaluesr,deltaAICr,AICweight)
 colnames(AICtable ) <- c("Model", "AIC","Delta AIC","AIC weight")
 print(AICtable)
```

The best model includes intensity and training, other possible models include all three covariates or just training (models with AIC weight > 0.1)

### Analysis of Deviance:
#### Level 1 Analysis of Deviance - compare to constant model
```{r,comapruing with constant model}
 Modellist <- list( )
 Modellist[[1]] <- modeli
 Modellist[[2]] <- modell
 Modellist[[3]] <- modelt
 modelcompare <- modelconst
 modelname <- c("intensity", "length","training") 
 # Create Analysis of Deviance
 dev <- rep(0,(length(Modellist)))
 for (i in 1:(length(Modellist))){dev[i]=Modellist[[i]]$deviance}
 devcompare <- modelcompare$deviance
 DF <- rep(0,(length(Modellist)))
 for (i in 1:(length(Modellist))){DF[i]=Modellist[[i]]$df.residual}
 DFcompare <- modelcompare$df.residual
 diffdev=devcompare-dev
 diffDF=DFcompare-DF
 pval=pchisq(diffdev, df=diffDF, lower.tail=FALSE)
 devtable = data.frame(modelname,diffdev,diffDF,pval)
 colnames(devtable) <- c("Model", "Diff Deviance","Diff DF","p-value")
 print(devtable)
```
Both intensity and training have p-values < 0.05, training has smallest p-values, so use to compare at level 2

#### Level 2 – compare to training
```{r,level2 comparing to training}
 Modellist <- list( )
 Modellist[[1]] <- modelit
 Modellist[[2]] <- modellt
 modelcompare <- modelt
 modelname <- c("intensity+training","length+training") 
 
 # Create Analysis of Deviance 
 dev <- rep(0,(length(Modellist)))
 for (i in 1:(length(Modellist))){dev[i]=Modellist[[i]]$deviance}
 devcompare <- modelcompare$deviance
 DF <- rep(0,(length(Modellist)))
 for (i in 1:(length(Modellist))){DF[i]=Modellist[[i]]$df.residual}
 DFcompare <- modelcompare$df.residual
 diffdev=devcompare-dev
 diffDF=DFcompare-DF
 pval=pchisq(diffdev, df=diffDF, lower.tail=FALSE)
 devtable = data.frame(modelname,diffdev,diffDF,pval)
 colnames(devtable) <- c("Model", "Diff Deviance","Diff DF","p-value")
 print(devtable)
```
Only adding intensity has a p-value < 0.05. Compare to intensity + training at the next level.


#### Level 3 Analysis of Deviance - compare to training+intensity
```{r, Level 3 compare to training+intensity}
 Modellist <- list( )
 Modellist[[1]] <- modelilt
 modelcompare <- modelit
 modelname <- c("intensity+length+training") 
 
 # Create Analysis of Deviance
 dev <- rep(0,(length(Modellist)))
 for (i in 1:(length(Modellist))){dev[i]=Modellist[[i]]$deviance}
 devcompare <- modelcompare$deviance
 DF <- rep(0,(length(Modellist)))
 for (i in 1:(length(Modellist))){DF[i]=Modellist[[i]]$df.residual}
 DFcompare <- modelcompare$df.residual
 diffdev=devcompare-dev
 diffDF=DFcompare-DF
 pval=pchisq(diffdev, df=diffDF, lower.tail=FALSE)
 devtable = data.frame(modelname,diffdev,diffDF,pval)
 colnames(devtable) <- c("Model", "Diff Deviance","Diff DF","p-value")
 print(devtable)
```
p-value > 0.05, so do not include length. Therefore, best model is intensity + training. This is the same as the best model according to AIC.


### Now consider adding random participant effects to this model:
#### Model with just staff:
```{r,model with just staff}
 modelits <- glmer(formula=complete~ intensity+factor(training)+(1|Staff),data = nmbdata,family = binomial())
```

#### Model with staff and participant:
```{r,Model with staff and participant}
 
 modelitsp <- glmer(formula=complete~ intensity+factor(training)+(1|Staff/Participant),data = nmbdata,family = binomial()) # ingnore the error boundary (singular) fit: see help('isSingular')
```

### LRT for whether need to add random staff effect
```{r,LRT for whether need to add random staff effect}
 LRT <- 2*(logLik(modelits)[1]-logLik(modelit)[1])
 pvalue <- 1-pchisq(LRT, df=1) 
 LRTresults <- data.frame(LRT,pvalue)
 print(LRTresults)
```
As the p-value > 0.05, this suggests there is no evidence of a random staff effect.

### AIC model selection:
```{r,AIC model selection}
 AIC(modelit,modelits,modelitsp)

```
According to AIC the best model has intensity and training but not a random staff or participant effect.

The LRT and AIC agree that the best model includes intensity and training.

### Results for best model:
```{r, Results for best model}
 summary(modelit)
```

In conclusion, the best model according to AIC includes both intensity and training, but not length and not a random staff nor participant effect. Other model selection (deviance and LRT) techniques agree with this conclusion.

The parameter estimates and their standard errors are as follows:
 

```{r, table}

Parameter	<-c('Intercept', 'Intensity','Training_1')
Estimate<-c(5.27, -0.83, 1.18)
Standard_Error<-c(2.20, 0.36, 0.43)
data.frame(Parameter,Estimate,Standard_Error)
````

As the intensity increases the probability of completing decreases. The participant is more likely to complete if they took part in training (1 represents took part in program). 


